Union and intersection contracts are hard, actually

### 2021-07-25 paper.tex
Revision: side-effects

Strikingly, we will see in Section~\ref{sec:issues-sem} that the introduction of
unions and intersections make (\ref{eq:cbn-expansion}) unsound \emph{even in a
pure setting}, preventing the kind optimization of
Figure~\ref{fig:optimizations-let-floating-ex} to fire in general.

>>>>

Higher-order contracts described in Section~\ref{sub-sec:contracts} already
introduce side-effects: semantically, the \nickel{blame} operator signaling
contract violation is equivalent to non-termination (error reporting aside).
But non-termination is a special kind of effect, if only because it preserves
referential transparency. In a lazy language like Nickel,
(\ref{eq:cbn-expansion}) still holds for all terms in the presence of
non-termination. In a strict language (call-by-value), (\ref{eq:cbn-expansion})
fails in all generality because the transformation may cause a previously
unevaluated term to be suddenly evaluated.  However, (\ref{eq:cbn-expansion})
holds given slight restrictions on the context $M$: for example, it holds as
long as at least one occurrence of $x$ in $M$ is in scope (i.e. not inside a
function). Beside, many languages out there already feature non-termination. For
such langauges, adding contracts does not change what optimizations are valid.

Strikingly, we will see in Section~\ref{sec:issues-sem} that the introduction of
unions and intersections \emph{breaks referential transparency}. Unions and
intersections make (\ref{eq:cbn-expansion}) unsound even in a lazy language,
preventing the kind optimization of
Figure~\ref{fig:optimizations-let-floating-ex} to fire in general. In a strict
language, the restrictions to (\ref{eq:cbn-expansion}) outlined above are not
sufficient anymore. If the effect of plain higher-order contracts behaves as
non-termination, the effect of union and intersection contracts rather behaves
as shared state.
